Supporting Information : The importance of a complete statistical description of dependence between variables for ecology and related fields

### 2018-10-10 Paper.Rmd
intro completed

Identifying and quantifying dependencies is the core of modelling in many applications, frequently found in finance and risk management. For many years, measuring and modeling dependencies has focused on the
Pearson correlation because of its ease in use. But correlation can provide little information about the underlying dependence structure in the case of asymmetric dependence because it is a measure of linear association between two variables. It may be better to use rank-based correlation measures such as Kendall’s tau and Spearman’s rho instead of Pearson correlation, when the marginal distribution of the variables are highly non-Gaussian. Copula, a ranked based dependence measurement between two variables with initially some non-uniformly distributed marginals, transforms the marginals into uniform distribution when joint distribution of ranked data is plotted. For example, in the first panel of Fig. \ref{fig_cop_pedag}, we showed the joint distribution of raw data (left) and copula data (right), where the marginals of raw data variables (x and y) were gamma distributed but in copula plot those marginals had uniform distribution. Copula plot was obtained by scaling individual ranked variables by a factor $(n+1)$ where $n$ was the number of sampling points, so that the copula was confined within a unit box. However, using a "copula-based approach" as a dependence measure reflects more knowledge about the dependence structure instead of summarizing in just a single number, the linear correlation coefficient. It tells us about the tail-dependence of the variables, i.e., the association between extreme values of two variables. If smaller values of each variable are more correlated to each other, then the corresponding copula structure is said to have "lower" tail-dependence as shown in the first panel of Fig. \ref{fig_cop_pedag} copula plot (right). Similarly, an "upper" tail-dependence copula indicates stronger correlation between higher extreme variables in the raw data. The flexibility of copulas allows us to model more complex dependence structures by isolating marginal from joint distribution structures, with the capability of modeling a rich variety of tail behavior, such as tail dependence versus tail independence, and reflection symmetry versus reflection asymmetry. Tail dependence occurs when the correlation between two variables increases as we get "further" in the tail (either or both) of the distribution. For example, in the upper panel of Fig. \ref{fig_cop_pedag}, different dependence structures, with same Kendall correlation $\tau = 0.5$, are shown which vary between dominantly lower tail dependent (lower tail - upper tail dependence being positive) from dominantly upper tail dependent (lower tail - upper tail dependence being negative) as we move along the row. Lower panel represents the similar scenario but with stronger correlation (Kendall's $\tau = 0.8$). To implement this, we need to select candidate bivariate copulas from many existing parametric copula families. <!--A very flexible bivariate copula can itself become a welcome candidate in the pool of bivariate copulas to be selected from.--> Here, throughout our studies for different datasets we used a universal pool of 16 bivariate copula families (having exclusively lower, exclusively upper, both or neither tail dependence), mentioned in the \nameref{Methods}.

Though a few examples have been found which shows copula approach to model dependencies in bioinformatics [@Kim2008], medical research [@Emura2016], a short period study of intersite dependence of rainfall data [@Serinaldi2008], a recent study on characterizing and assessing risks in drought management [@She2018] or monitoring changes in precipitation-extremes [@Goswami2018], the multivariate copula approach still, we believe, remains unexplored in ecology or related field and thus have potential scope to better capture the evolution of ecological time series data. In ecological perspective, when we find a dominantly lower (upper) tail dependence structure between two species abundance timeseries, it means both incidence are rare (abundant) synchronously. If one variable is an environmental variable and another is an observable (like abundance of a species) then it can also explain the causal relationship between them. Here, we analyzed the dependence structure for several bivariate and multivariate ecological, community and biogeochemical datasets. In addition to a copula based \nameref{Model_selection}, we also carried out a \nameref{NPA} with three different statistics to verify our model selection based results.

Basically, we want to address the following questions. \textbf{(Q1)} Do datasets in ecology and related fields have non-Normal copula srtructure? Do they show tail dependence distinct from that of a Normal copula, and in particular do they show asymmetric tail dependence? \textbf{(Q2)} If so, what are some possible causes of non-Normal copula structure? And \textbf{(Q3)} what are the consequences of non-Normal copula structure and tail dependence for our understanding of biology and for applications? 
In \nameref{Material&Methods}, we described briefly about each bivariate and multivariate datasets we used to answer these questions and provided a detailed descriptions of implemented methods. In the \nameref{Results} section, we summarized our finidings for each datasets. Next two sections and are focused on possible causes and consequences of non-Normal copula structure in ecology and related field. Finally, we concluded in the \nameref{Discussion} section.

\begin{figure}[!h]
\begin{center}
\includegraphics[width=4 cm,height=4 cm]{./Results/rawdata.pdf}
\includegraphics[width=4 cm,height=4 cm]{./Results/copdata.pdf}\\
\includegraphics[width=16 cm,height=8 cm]{./Results/copula_pedagog_fig_BB1.pdf}
\caption{Pedagogical figure showing BB1 copula with varying tail-dep difference but of same Kendall's tau along each row and with same tail-dep difference but of different Kendall's tau along each column.\label{fig_cop_pedag}}
\end{center}
\end{figure}

# Material and methods \label{Material&Methods}

@@

Considered datasets ranged widely from population ecology, community ecology , evolutionary and biogeochemical sphere. 
We used four datasets in each category of bivariate (soil C and N data, BMR data, Anolis data, Ceder-creek experiment data) and multivariate (aphid-abundance data, aphid-firstflight data, plankton-abundance data and methane-flux data) examples. Details of each dataset used were given in \ref{SM-Data_detailed}.

@@

```{r tab_bivar_summary, echo=F, results='markup'}

@@

We also showed **(ii)** asymmetric copula structure affected slope of spatial Taylor's Law differently than as expected from symmetric Normal copula (For details see Appendix \ref{SM-Consequence_TL}). For example, our model simulation showed lower tail dependent Clayton copula had much higher slope (spatial-Taylor's law) than symmetric Normal copula, which in turn had higher slope than upper tail dependent Survival Clayton family (Fig. \ref{SM-fig_spTL}(I)). Finally, we studied **(iii)** the effects of environmental noise, having a non-Normal copula, on extinction risk (calculated after 25 years) with two metapopulation model, namely, Lewton-Cohen model and Ricker model (For details see Appendix \ref{SM-Consequence_ext_risk}). We considered patches were arranged in a linear chain fashion and the population may disperse to its nearerst two neighbors (local dispersion) or to any patches (global dispersion). For both model with both dispersion, increasing number of habitat patches and increasing carrying capacity, separately, had decreased extinction risk of the populations which was expected (Fig. \ref{SM-fig_risk_numlocs_effect_local_disp} - Fig. \ref{SM-fig_risk_K_effect}). For Lewton-Cohen model, we found greater extinction risk if populations were affected by left tail dependence environmental noise copula than right tail dependent one. For Ricker model, we got the similar observation for smaller values of $r$, the intrinsic growth rate, but for higher values of $r$, there was a turn-over in extinction risks, and right tail dependent noise copula caused population go extinct faster.

>>>>

Identifying and quantifying dependencies is the core of modelling in many applications, frequently found in finance and risk management. For many years, measuring and modeling dependencies has focused on the
Pearson correlation because of its ease in use. But correlation can provide little information about the underlying dependence structure in the case of asymmetric dependence because it is a measure of linear association between two variables. It may be better to use rank-based correlation measures such as Kendall’s tau and Spearman’s rho instead of Pearson correlation, when the marginal distribution of the variables are highly non-Gaussian. Copula [@joe2014_dependence; @nelsen2007_copula], a ranked based dependence measurement between two variables with initially some non-uniformly distributed marginals, transforms the marginals into uniform distribution when joint distribution of ranked data is plotted. For example, in the first panel of Fig. \ref{fig_cop_pedag}, we showed the joint distribution of raw data (left) and copula data (right), where the marginals of raw data variables (x and y) were gamma distributed but in copula plot those marginals had uniform distribution. Copula plot was obtained by scaling individual ranked variables by a factor $(n+1)$ where $n$ was the number of sampling points, so that the copula was confined within a unit box. However, using a "copula-based approach" as a dependence measure reflects more knowledge about the dependence structure instead of summarizing in just a single number, the linear correlation coefficient. It tells us about the tail-dependence of the variables, i.e., the association between extreme values of two variables. If smaller values of each variable are more correlated to each other, then the corresponding copula structure is said to have "lower" tail-dependence as shown in the first panel of Fig. \ref{fig_cop_pedag} copula plot (right). Similarly, an "upper" tail-dependence copula indicates stronger correlation between higher extreme variables in the raw data. Tail dependence occurs when the correlation between two variables increases as we get "further" in the tail (either or both) of the distribution. For example, in the middle panel of Fig. \ref{fig_cop_pedag}, different dependence structures, with same Kendall correlation $\tau = 0.5$, are shown which vary between dominantly upper tail dependent (lower tail - upper tail dependence being negative) to dominantly lower tail dependent (lower tail - upper tail dependence being positive) as we move along the row. Lower panel represents the similar scenario but with stronger correlation (Kendall's $\tau = 0.8$). So, the flexibility of copulas allows us to model more complex multivariate dependence structures by isolating marginal from joint distribution structures, with the capability of modeling a rich variety of tail behavior, such as tail dependence versus tail independence, and reflection symmetry versus reflection asymmetry. To implement this, we need to select candidate bivariate copulas from many existing parametric copula families. <!--A very flexible bivariate copula can itself become a welcome candidate in the pool of bivariate copulas to be selected from.--> Here, throughout our studies for different datasets we used a universal pool of 16 bivariate copula families (having exclusively lower, exclusively upper, both or neither tail dependence), mentioned in the \nameref{Methods}.

Though a few examples have been found which shows copula approach to model dependencies in bioinformatics [@Kim2008], medical research [@Emura2016], a short period study of intersite dependence of rainfall data [@Serinaldi2008], a recent study on characterizing and assessing risks in drought management [@She2018] or monitoring changes in precipitation-extremes [@Goswami2018], the multivariate copula approach still, we believe, remains unexplored in ecology or related field and thus have potential scope to better capture the evolution of ecological time series data. In ecological perspective, when we find a dominantly lower (upper) tail dependence structure between two species abundance timeseries, it means both incidence are rare (abundant) synchronously. If one variable is an environmental variable and another is an observable (like abundance of a species) then it can also explain the causal relationship between them. On a similar note, if a species' abundance time series from different sampling locations shows multivariate dependence structure, then we can analyze them as pairwise bivariate copula and find if there is any tail dependence or not. If those copula came up with asymmetric tail dependence, then it can be viewed as a measure of "spatial synchrony" [@Liebhold2004; @walteretal2017], which is a celebrated phenomenon in population ecology. In this article, we analyzed the dependence structure for several bivariate and multivariate ecological, community and biogeochemical datasets. In addition to a copula based \nameref{Model_selection}, we also carried out a \nameref{NPA} with three different statistics to verify our model selection based results.

We want to address the following questions. \textbf{(Q1)} Do datasets in ecology and related fields have non-Normal copula srtructure? Do they show tail dependence distinct from that of a Normal copula, and in particular do they show asymmetric tail dependence? \textbf{(Q2)} If so, what are some possible causes of non-Normal copula structure? And \textbf{(Q3)} what are the consequences of non-Normal copula structure and tail dependence for our understanding of biology and for applications? 
In \nameref{Data}, we described briefly about each bivariate and multivariate datasets we used to answer these questions and provided a detailed descriptions of implemented methods in \nameref{Methods}. In the \nameref{Results} section, we summarized our finidings for each datasets. Next two sections are focused on possible causes and consequences of non-Normal copula structure in ecology and related field with applications shown in real data as well as some simulations. Finally, we concluded in the \nameref{Discussion} section.

\begin{figure}[!h]
\begin{center}
\includegraphics[width=4 cm,height=4 cm]{./Results/rawdata.pdf}
\includegraphics[width=4 cm,height=4 cm]{./Results/copdata.pdf}\\
\includegraphics[width=16 cm,height=8 cm]{./Results/copula_pedagog_fig_BB1.pdf}
\caption{First panel : raw data plot has non-uniform (gamma distributed) marginal whereas copula makes its marginal uniform and indicating lower tail dependence in joint distribution, Second and third panel : pedagogical figure showing BB1 copula with varying tail-dependence difference but of same Kendall's tau along each row and with same tail-dependence difference but of different Kendall's tau along each column.\label{fig_cop_pedag}}
\end{center}
\end{figure}

<!--# Material and methods \label{Material&Methods}-->

@@

<!--
Considered datasets ranged widely from population ecology, community ecology , evolutionary and biogeochemical sphere. 
We used four datasets in each category of bivariate (soil C and N data, BMR data, Anolis data, Ceder-creek experiment data) and multivariate (aphid-abundance data, aphid-firstflight data, plankton-abundance data and methane-flux data) examples. Details of each dataset used were given in \ref{SM-Data_detailed}.-->

For bivariate analysis we used four datasets which are briefly described as follows;
\textbf{(i)} to explore the dependence of **soil organic C** and total **soil N** across a wide geographic range (conterminous US), ecosystem type, and traceable methodology we used the RaCA (Rapid Carbon Assessment)
database [@Wills2014, **or Wills and Loecke pending???**]. Specifically, the soil organic C and total soil N
stocks (Mg C or N m$^{-2}$ ) to 1m of soil depth from the central pedon of the 5930 sites were included in this
study. Then, we used extensive set of basal metabolic rate data for \textbf{(ii)} birds [@Mcnab_2009] and \textbf{(iii)} mammals [@Mcnab_2008] as our second and third bivariate datasets to explore the dependence between log(BMR) and log(mass) of respective species. \textbf{(iv)} We used [plant above ground biomass](https://www.cedarcreek.umn.edu/research/data/dataset?ple120) and [plant species percent cover](https://www.cedarcreek.umn.edu/research/data/dataset?pce120) data from a biodiversity experiment [@Tilman2001; @Tilman2006] to calculate plant-aboveground biomass and Shannon's diversity index (H), averaged over each plot (9m by 9m), for year 2000.

Similarly, we applied multivariate analysis on four datasets; first two data sets [for aphid-count and aphid-firstflight date] were extracted from the Rothamsted Insect Survey suction-trap data [@Macaulay1988, @Harrington2014]. These two annual timesries (1976-2010) measurements are from 11 locations acroos UK for 20 aphid species. We used the same data format as used by @Sheppard2016 for their analysis on the impact of climate on aphids' spatial synchrony. In our case, we chose species=10 [Name : "Green spruce aphid"] for aphid-abundance data and species=11 [Name : "Leaf-curling plum aphid"] for aphid-firstflight data and considered only those locations which had atleast 30 years of data (see Figure \ref{SM-fig_aphid_map}).

Our third multivariate dataset is a plankton-count data collected at 26 locations over the areas of the North Sea and British seas for 22 species and for 56 years (1958-2013). These data were obtained from the Continuous Plankton Recorder (CPR) data set and maintained by the Sir Alistair Hardy Foundation for Ocean Science (SAHFOS).
<!--The CPR is a device which is towed across the ocean behind a ship, filtering plankton from seawater with a continuously
unwinding silk ribbon. The ribbon was cut into 4 inch segments, each corresponding to a particular time and
location in the ocean. Organisms were identified and counted (possibly with sub-sampling) microscopically [@Colebrook1979]. All the samples available for a given variable in a given area in a given month of a given year were
averaged, producing a monthly time series for each year. Where data for a given month was unavailable the missing value was replaced with a median value for that month (drawn from all years in which data was available).--> [@Sheppard2017] converted monthly values to yearly values by averaging over all months. But here we only allowed maximum 4 months in a year with missing data which were replaced by median value for that month (drawn from all years in which data was available). We used data for species=16 [Name : "Ceratium furca"] from those locations which had atleast 45 years time series data (see Figure \ref{SM-fig_plankton_map}).

Finally, methane ($CH_{4}$) fluxes between the soil or water surface and the troposphere were measured at the
Great Miami Wetland Mitigation Bank ($36^{\circ}46^{'}51^{''}$ N, $84^{\circ}20^{'}26^{''}$ W) in Trotwood, Montgomery County, Ohio. Additional information on the history of the site and soil types is detailed in @Jarecke2016. Briefly, the closed static flux chamber method was used [@Holland1999] to measure $CH_{4}$ fluxes at 28 locations across a hydrologic gradient (upland to seasonal wetland) at daily to weekly interval from September 2015 to September 2016. **See and Smyth et al, pending for
additional details???** on greenhouse gas flux methods for this site. We threw out all locations that had fewer than 50 dates data.

@@

\newpage
```{r tab_bivar_summary, echo=F, results='markup'}

@@

We also showed **(ii)** asymmetric copula structure affected slope of spatial Taylor's Law differently than as expected from symmetric Normal copula (For details see Appendix \ref{SM-Consequence_TL_simulation}). For example, our model simulation showed lower tail dependent Clayton copula had much higher slope (spatial-Taylor's law) than symmetric Normal copula, which in turn had higher slope than upper tail dependent Survival Clayton family (Fig. \ref{SM-fig_spTL}(I)). Finally, we studied **(iii)** the effects of environmental noise, having a non-Normal copula, on extinction risk (calculated after 25 years) with two metapopulation model, namely, Lewton-Cohen model and Ricker model (For details see Appendix \ref{SM-Consequence_ext_risk}). We considered patches were arranged in a linear chain fashion and the population may disperse to its nearerst two neighbors (local dispersion) or to any patches (global dispersion). For both model with both dispersion, increasing number of habitat patches and increasing carrying capacity, separately, had decreased extinction risk of the populations which was expected (Fig. \ref{SM-fig_risk_numlocs_effect_local_disp} - Fig. \ref{SM-fig_risk_K_effect}). For Lewton-Cohen model, we found greater extinction risk if populations were affected by left tail dependence environmental noise copula than right tail dependent one. For Ricker model, we got the similar observation for smaller values of $r$, the intrinsic growth rate, but for higher values of $r$, there was a turn-over in extinction risks, and right tail dependent noise copula caused population go extinct faster.

### 2018-10-10 SupportingInformation.Rmd
intro completed

@@

<!--
## Methane-flux data  \label{Methane_data}

Methane ($CH_{4}$) fluxes between the soil or water surface and the troposphere were measured at the
Great Miami Wetland Mitigation Bank ($36^{\circ}46^{'}51^{''}$ N, $84^{\circ}20^{'}26^{''}$ W) in Trotwood, Montgomery County, Ohio. Additional information on the history of the site and soil types is detailed in @Jarecke2016. Briefly, the closed static flux chamber method was used [@Holland1999] to measure $CH_{4}$ fluxes at 28 locations across a hydrologic gradient (upland to seasonal wetland) at
daily to weekly interval from September 2015 to September 2016. **See and Smyth et al, pending for
additional details???** on greenhouse gas flux methods for this site.

* We threw out all locations that had fewer than 50 dates data.-->

@@

# Description of Data \label{Data_detailed}

@@

<!--The raw data and copula for each four datasets are shown along each column of Figure \ref{MT-fig_biv_raw_cop}.-->

@@

<!--Figure \ref{MT-fig_aphid_raw_cop} represents raw and copula data plot for aphid count ("Green spruce aphid") and aphid-firstflight date ("Leaf-curling plum aphid") taken from two locations.-->

@@

Finally, methane ($CH_{4}$) fluxes between the soil or water surface and the troposphere were measured at the
Great Miami Wetland Mitigation Bank ($36^{\circ}46^{'}51^{''}$ N, $84^{\circ}20^{'}26^{''}$ W) in Trotwood, Montgomery County, Ohio. Additional information on the history of the site and soil types is detailed in @Jarecke2016. Briefly, the closed static flux chamber method was used [@Holland1999] to measure $CH_{4}$ fluxes at 28 locations across a hydrologic gradient (upland to seasonal wetland) at
daily to weekly interval from September 2015 to September 2016. **See and Smyth et al, pending for
additional details???** on greenhouse gas flux methods for this site. We threw out all locations that had fewer than 50 dates data.

@@

```{r tab_soilCN_asym,echo=F,results="markup"}
#knitr::opts_knit$set(kable.force.latex = TRUE)
tab_soilCN_asym<-make_table_stat_fn(data=stat_soilCN_ga)
knitr::kable(tab_soilCN_asym, booktabs = T, 
             format = "pandoc",
      #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
      caption = "**Soil CN data set results** for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis. Three statistics are listed in the left column that measured asymmetry of tail dependence by comparing the distribution ranges $0-0.1$ and $0.9-1$ (see \\nameref{MT-NPA}). One thousand Kendall-preserving and 1000 Spearman-preserving surrogates with normal copula structure were created, and the same statistics were computed for the surrogates. A table entry $<N$ indicates the value of the given statistic on the data was less than its value on $N$ of the surrogates, so entries of the form $<X$ for $X$ equal to $975$ or above indicate that upper-tail dependence was significantly stronger than lower-tail dependence.\\label{tab_soilCN_asym}")
#kable_styling(, position = "float_left")
```

@@

}
#----------------------------------
#calling the above function on 4 multivariable datasets
m1<-multivar_tab_summary_fn(r=RES_aphid_count,rs=stat_aphid_count)
m2<-multivar_tab_summary_fn(r=RES_aphid_ff,rs=stat_aphid_ff)
m3<-multivar_tab_summary_fn(r=RES_plankton_north_sea,rs=stat_plankton_north_sea)
m4<-multivar_tab_summary_fn(r=RES_methane,rs=stat_methane)
tab_multivar_summary<-cbind(m1,m2,m3,m4)
colnames(tab_multivar_summary)<-c("Aphid_abundance","Aphid_firstflight","Plankton","Methane-flux")
saveRDS(tab_multivar_summary,"./Results/tab_multivar_summary.RDS")
```

@@

\newpage
```{r tab_stat_npa_asym_4,echo=F,results="markup"}
temp<-make_table_stat_fn(data=tab_asym[[4]])
knitr::kable(temp, booktabs = T, 
             format = "pandoc",
      #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
      caption =paste(cnm[4]," : A table entry $<N$(or $>N$) indicates the value of the given statistic on the data was less (or more) than its value on $N$ of the surrogates, so entries of the form $<X$ (or $>X$) for $X$ equal to $975$ or above indicate that upper-tail (or lower-tail) dependence was significantly stronger than lower-tail dependence.\\label{tab_stat_npa_asym_4}"))
```

@@

\newpage

#  References
<div id="refs"></div>

>>>>

```{r cop_fig_intro,echo=F,results="hide",cache=T, cache.extra=list(seed,mtime("copsurrognd.R"))}
library(copula)
library(VineCopula)
library(mvtnorm)
set.seed(seed)

source("copsurrognd.R")

#x <- read.table("./Data/dataset/x.txt")
#y <- read.table("./Data/dataset/y.txt")
#mat<-cbind(x,y)
#colnames(mat)<-c("x","y")

#***constants
N<-200 #length of time series
n<-2 #number of sampling locations
numdat<-1 #the number of datasets to get from the normal copula
rho<-0.8 #normal-copula parameter for dependence between locations
#this is a covariance, see below call to rmvnorm

#***generate the starting data, which has normal copula structure
#because it comes originally from a multivariate normal, before 
#transformation
sig<-matrix(rho,n,n)
diag(sig)<-1
m<-rmvnorm(N*numdat,mean=rep(0,n),sigma=sig)
m<-array(m,c(N,numdat,n))
m<-aperm(m,c(1,3,2)) # 2nd dim is normally distributed e.g. hist(m[,1,])
gshape<-7.5
gscale<-1 #gamma parameters

cop_to_marg_trans<-function(x){
  return(qgamma(x,shape=gshape,scale=gscale)) #this is where we choose the marginals,
  #should be something skewed
}
# pnorm(m) makes normal to uniform
# then qgamma(...) makes uniform to gamma distributed
m<-cop_to_marg_trans(pnorm(m))

ccop<-claytonCopula(2,2)
csurrog<-array(NA,dim(m))
csurrog[,,1]<-copsurrognd(m[,,1],targetcop=ccop,numsurrog=1)
mat<-csurrog[,,1]

pdf("./Results/rawdata.pdf",height = 4, width = 4)
scale<-2
xhist <- hist(mat[,1], breaks=30, plot=FALSE)
yhist <- hist(mat[,2], breaks=30, plot=FALSE) 
top <- max(c(xhist$counts, yhist$counts)) 
xrange <- range(mat[,1])
yrange <- range(mat[,2])
nf <- layout(matrix(c(2,0,1,3),2,2,byrow=TRUE), c(4,1), c(1,4), TRUE) 
op<-par(mar=c(3,3,1,1),mgp=c(2, 1, 0))  
plot(mat[,1], mat[,2], xlim=xrange, ylim=yrange, xlab="x", ylab="y", cex.lab=1.5)
legend("bottomright","raw data plot",bty="n",cex=1.5,text.col="magenta")
op<-par(mar=c(0,3,1,1)) 
barplot(xhist$counts/scale, axes=FALSE, ylim=c(0, top), space=0) 
op<-par(mar=c(3,0,1,1)) 
barplot(yhist$counts/scale, axes=FALSE, xlim=c(0, top), space=0, horiz=TRUE)
par(op)
dev.off()

pdf("./Results/copdata.pdf",height = 4, width = 4)
copmat<-pobs(mat)
xhist <- hist(copmat[,1], breaks=30, plot=FALSE)
yhist <- hist(copmat[,2], breaks=30, plot=FALSE) 
top <- max(c(xhist$counts, yhist$counts)) 
xrange <- c(0,1)
yrange <- c(0,1)
nf <- layout(matrix(c(2,0,1,3),2,2,byrow=TRUE), c(4,1), c(1,4), TRUE) 
op<-par(mar=c(3,3,1,1),mgp=c(2, 1, 0)) 
plot(copmat[,1], copmat[,2], xlim=xrange, ylim=yrange, xlab="u=rank(x)/(n+1)", ylab="v=rank(y)/(n+1)",cex.lab=1.5)
legend("bottomright","copula plot",bty="n",cex=1.5,text.col="magenta")
op<-par(mar=c(0,3,1,1)) 
barplot(xhist$counts/scale, axes=FALSE, ylim=c(0, top), space=0) 
op<-par(mar=c(3,0,1,1)) 
barplot(yhist$counts/scale, axes=FALSE, xlim=c(0, top), space=0, horiz=TRUE)
par(op)
dev.off()
```

@@

@@

<!--# Description of Data \label{Data_detailed}

@@

The raw data and copula for each four datasets are shown along each column of Figure \ref{MT-fig_biv_raw_cop}.

@@

Figure \ref{MT-fig_aphid_raw_cop} represents raw and copula data plot for aphid count ("Green spruce aphid") and aphid-firstflight date ("Leaf-curling plum aphid") taken from two locations.

@@

Finally, methane ($CH_{4}$) fluxes between the soil or water surface and the troposphere were measured at the
Great Miami Wetland Mitigation Bank ($36^{\circ}46^{'}51^{''}$ N, $84^{\circ}20^{'}26^{''}$ W) in Trotwood, Montgomery County, Ohio. Additional information on the history of the site and soil types is detailed in @Jarecke2016. Briefly, the closed static flux chamber method was used [@Holland1999] to measure $CH_{4}$ fluxes at 28 locations across a hydrologic gradient (upland to seasonal wetland) at
daily to weekly interval from September 2015 to September 2016. **See and Smyth et al, pending for
additional details???** on greenhouse gas flux methods for this site. We threw out all locations that had fewer than 50 dates data.-->

@@

\newpage
```{r tab_soilCN_asym,echo=F,results="markup"}
#knitr::opts_knit$set(kable.force.latex = TRUE)
tab_soilCN_asym<-make_table_stat_fn(data=stat_soilCN_ga)
knitr::kable(tab_soilCN_asym, booktabs = T, 
             format = "pandoc",
      #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
      caption = "**Soil CN data set results** for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis. Three statistics are listed in the left column that measured asymmetry of tail dependence by comparing the distribution ranges $0-0.1$ and $0.9-1$ (see \\nameref{MT-NPA}). One thousand Kendall-preserving and 1000 Spearman-preserving surrogates with normal copula structure were created, and the same statistics were computed for the surrogates. A table entry $<N$ indicates the value of the given statistic on the data was less than its value on $N$ of the surrogates, so entries of the form $<X$ for $X$ equal to $975$ or above indicate that upper-tail dependence was significantly stronger than lower-tail dependence.\\label{tab_soilCN_asym}")
#kable_styling(, position = "float_left")
```

@@

}
#----------------------------------
#calling the above function on 4 multivariable datasets
m1<-multivar_tab_summary_fn(r=RES_aphid_count,rs=stat_aphid_count)
m2<-multivar_tab_summary_fn(r=RES_aphid_ff,rs=stat_aphid_ff)
m3<-multivar_tab_summary_fn(r=RES_plankton_north_sea,rs=stat_plankton_north_sea)
m4<-multivar_tab_summary_fn(r=RES_methane,rs=stat_methane)
tab_multivar_summary<-cbind(m1,m2,m3,m4)
colnames(tab_multivar_summary)<-c("Green-spruce aphid_count","Leaf-plum curling aphid_firstflight","C. furca_count","Methane-flux")
saveRDS(tab_multivar_summary,"./Results/tab_multivar_summary.RDS")
```

@@

```{r tab_stat_npa_asym_4,echo=F,results="markup"}
temp<-make_table_stat_fn(data=tab_asym[[4]])
knitr::kable(temp, booktabs = T, 
             format = "pandoc",
      #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
      caption =paste(cnm[4]," : A table entry $<N$(or $>N$) indicates the value of the given statistic on the data was less (or more) than its value on $N$ of the surrogates, so entries of the form $<X$ (or $>X$) for $X$ equal to $975$ or above indicate that upper-tail (or lower-tail) dependence was significantly stronger than lower-tail dependence.\\label{tab_stat_npa_asym_4}"))
```

@@

\clearpage

# References

<!--
<div id="refs"></div>
-->
